 

### 2023-02-10 paper1.Rmd
Clean up of formatting and typesetting throughout.

---
title: ''
output: pdf_document
header-includes:
  - \usepackage{float}

---

# Chapter 1: Quantifying the global burden of mental disorders and their economic value

>>>>

```{r paper1code_libraries_data}

# Loading libraries
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(sp)
library(broom)
library(sf)
library(scales)
library(viridis)
library(data.table)
library(ggpubr)
library(lemon)
library(rnaturalearth)
library(rnaturalearthdata)
library(sysfonts)
library(showtext)
font_add_google("Source Sans Pro")
library(gridExtra)
showtext_auto() 
showtext_opts(dpi = 300)

# Setting directories - change these to your local directory
path <- "C:/Users/danie/Dropbox (Harvard University)/Important Files/3. PhD/WQE/WQE II/gmh_econ/"
datapath <- paste(path,'data', sep = "/")
resultspath <- paste(path,'results', sep = "/")

######################
##       DATA       ##
######################

# Loading IHME Global Burden of Disease Data

# The data are available for download at http://ghdx.healthdata.org/gbd-results-tool
# Data should be downloaded for DALYs, YLLs, YLDs, deaths, and prevalence by cause and geographies
# Save the file as IHME-GBD_2019_DATA-2019.csv
# If prevalence data are saved separately, save it as IHME-GBD_2019_DATA-2019-prev.csv

data <- read.csv(file = file.path(datapath,"IHME-GBD_2019_DATA-2019.csv"))

# Adding alpha-3 codes from ISO 3166-1 (i.e., World Bank Codes)
ihme_crosswalk <- read_excel(path = file.path(datapath, "ihme_crosswalk.xlsx")) %>% 
  select(-c("location_name"))
data <- left_join(data, ihme_crosswalk, by = "location_id")

# Adding World Bank income classifications
income_level <- read_excel(path = file.path(datapath, "OGHIST.xls"), sheet = 2) %>% 
  select(-(c(2:34))) %>% 
  rename("income_level" = "2019")
data <- left_join(data, income_level, by = "iso_code")

# Adding population (backed out using IHME rate data for consistency)
data <- data %>% unique() %>% gather("numeric_name", "est", "val":"lower") 
data_number <- data %>% filter(metric_id == 1) %>% rename("number" = "est")
data_rate <- data %>% filter(metric_id == 3) %>% select("measure_id","location_id", "sex_id", "age_id", "cause_id", "year", "numeric_name", "est") %>% rename("rate" = "est")
data <- left_join(data_number, data_rate, by = c("measure_id","location_id", "sex_id", "age_id", "cause_id", "year", "numeric_name")) %>% 
  select(!c("metric_id", "metric_name")) %>%
  unique()
data$population_100k <- data$number/data$rate
data$population <- data$population_100k  * 100000

# Removing unused columns
data <- data %>% select(!c("rate"))  %>% 
  select(!c("sex_id", "sex_name", "age_id", "age_name", "year")) %>%
  filter(measure_id != 5)

#######################################################
##  ORIGINAL AND 2016 REALLOCATION APPROACH  ##
##################################################

# VIGO EL AL. 2016 WEIGHTS: ALLOCATION OF THE FOLLOWING TO MENTAL HEALTH 
# (PERCENTS DENOTE AMOUNT REALLOCATED TO MENTAL HEALTH)

# Dementia - 100%
# Epilepsy - 100%
# Migraine - 100%
# Tension-type headache - 100%
# Self-harm - 100%
# Chronic pain syndrome currently attributed to musculoskeletal disorders - 33%

# Naming relevant IHME cause IDs
mental_disorder_cause_id <- 558
dementia_cause_id1 <- 543 # Alzheimer's disease and other dementias
dementia_cause_id2 <- 544 # Parkinson's disease
epilepsy_cause_id <- 545 # Idiopathic epilepsy
migraine_cause_id <- 547
tension_type_headache_cause_id <- 548
self_harm_cause_id <- 718
alcohol_use_disorders_cause_id <- 560
drug_use_disorders_cause_id <- 561
low_back_pain_cause_id <- 630
neck_pain_cause_id <- 631
other_musculoskeletal_cause_id <- 639
all_cause_id <- 294

# Original allocation

# Cause ID 558 reflects the GBD 2019 mental disorder allocation
# These lines of code retain this allocation and save it under "original value"
cause_id <- c(mental_disorder_cause_id)
original_weights <- c(1)
original_weights_df <- data.frame(cause_id, original_weights)
data_rev <- left_join(data, original_weights_df, by = c("cause_id"))
data_rev$original_weights[is.na(data_rev$original_weights)] <- 0
data_rev$original_value <- data_rev$number * data_rev$original_weights

# 2016 reallocation

cause_id <- c(mental_disorder_cause_id,
              dementia_cause_id1,
              dementia_cause_id2,
              epilepsy_cause_id,
              migraine_cause_id,
              tension_type_headache_cause_id,
              self_harm_cause_id,
              alcohol_use_disorders_cause_id,
              drug_use_disorders_cause_id,
              low_back_pain_cause_id,
              neck_pain_cause_id,
              other_musculoskeletal_cause_id,
              mental_disorder_cause_id,
              dementia_cause_id1,
              dementia_cause_id2,
              epilepsy_cause_id,
              migraine_cause_id,
              tension_type_headache_cause_id,
              self_harm_cause_id,
              alcohol_use_disorders_cause_id,
              drug_use_disorders_cause_id,
              low_back_pain_cause_id,
              neck_pain_cause_id,
              other_musculoskeletal_cause_id,
              mental_disorder_cause_id,
              dementia_cause_id1,
              dementia_cause_id2,
              epilepsy_cause_id,
              migraine_cause_id,
              tension_type_headache_cause_id,
              self_harm_cause_id,
              alcohol_use_disorders_cause_id,
              drug_use_disorders_cause_id,
              low_back_pain_cause_id,
              neck_pain_cause_id,
              other_musculoskeletal_cause_id)

numeric_name <- c("val", "val", "val", "val", "val","val", "val", "val", "val", "val", "val","val",
                  "lower","lower","lower","lower","lower","lower","lower","lower","lower","lower","lower","lower",
                  "upper","upper","upper","upper","upper","upper","upper","upper","upper","upper","upper","upper")

# Allocations of low back pain, neck pain, and other musculoskeletal disorders is based on
# the following section of Vigo et al. 2016

# "Considering that a fraction of low back, neck pain,
# and 50% of other musculoskeletal pain potentially
# corresponds to chronic pain syndromes, and for the
# purposes of producing a more accurate estimation and
# stimulating debate, we assume given the limited data
# that one third (rather than zero percent, as it is now) of
# the disease burden of these pain syndromes is
# potentially attributable to mental disorders and explore
#  the effect on mental illness burden calculations..."

rev_2016_weights <- c(1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      2/6, # Allocating 1/3
                      2/6,
                      1/2*2/6,

                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1/6, # Allocating 1/6
                      1/6,
                      1/2*1/6,

                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      1,
                      3/6, # Allocating 1/2 
                      3/6,
                      1/2*3/6)

rev_2016_weights_df <- data.frame(cause_id,numeric_name, rev_2016_weights)
data_rev <- left_join(data_rev, rev_2016_weights_df, by = c("cause_id", "numeric_name"))

data_rev$rev_2016_weights[is.na(data_rev$rev_2016_weights)] <- 0
data_rev$rev_2016_value <- data_rev$number * data_rev$rev_2016_weights

############################
##  COMPOSITE APPROACH  ##
############################

# Due to the distributive property of the Population Attributable Fraction, a category-specific attribution fraction is estimated
# using the formula from Rockhill et al, 1998:https://pubmed.ncbi.nlm.nih.gov/9584027/

natural_cause_id <- 409 #NCDs
unnatural_cause_id <- 687 # Injuries

cause_id <- c(natural_cause_id,
              unnatural_cause_id,
              natural_cause_id,
              unnatural_cause_id,
              natural_cause_id,
              unnatural_cause_id)

numeric_name <- c("val", "val", 
                  "lower","lower",
                  "upper","upper")

relative_risk <- c(1.8,
                   7.22,
                   1.71,
                   6.43,
                   1.88,
                   8.12)

relative_risk_df <- data.frame(cause_id,numeric_name, relative_risk)

data_rev <- left_join(data_rev, relative_risk_df, by = c("cause_id", "numeric_name"))

data_prev <- read.csv(file = file.path(datapath,"IHME-GBD_2019_DATA-2019-prev.csv"))
data_prev <- data_prev %>%
  gather("numeric_name", "prevalence", "val":"lower")  %>% 
  select(!c("sex_id", "sex_name", "age_id", "age_name", "year")) %>% filter(cause_id == mental_disorder_cause_id) %>% select("location_id", "numeric_name", "prevalence")
data_rev  <- left_join(data_rev, data_prev, by = c("location_id", "numeric_name"))

# Two approaches to calculating the PAF are presented here. Later, only one will be retained for the creation of graphs, tables, and charts.
# In Walker et al., the PAF formula numbered as formula 4 in Rockhill et al. is used, in which prevalence is meant to be the
# prevalence of mental disorders among cases (i.e., the deceased). The conventional formula, formula 2, takes prevalence as 
# the prevalence of mental disorders in the general population, but is not internally valid in cases of confounding.

data_rev$paf2 <-     (data_rev$prevalence  * ((data_rev$relative_risk - 1))) /   ((data_rev$prevalence  * ((data_rev$relative_risk - 1))+1))
data_rev$paf4 <-     data_rev$prevalence  * ((data_rev$relative_risk - 1) /    data_rev$relative_risk)

data_rev <- data_rev%>% gather("paf_formula", "paf", "paf2":"paf4") 
data_rev$paf[is.na(data_rev$paf)] <- 0
data_rev$rev_2015_value <- data_rev$number * data_rev$paf

# Combine original and re-allocated values by location, measure, PAF formula used, and numeric value (point estimate, and upper or lower bounds)
data_rev <- data_rev %>% 
  group_by(location_id, measure_id, numeric_name, paf_formula) %>% 
  mutate(rev_2016_value_agg = sum(rev_2016_value),  rev_2015_value_agg = sum(rev_2015_value),original_value_agg = sum(original_value)) %>%
  rename("measure_total" = number) %>%
  ungroup %>%
  filter(cause_id == "294") %>% select(!c("original_weights":"relative_risk", "paf":"rev_2015_value"))

# Clean up

############################
##  COMPOSITE METHOD  ##
############################

# This method pulls YLLs from Walker et al. and YLDs from Vigo et al. together to re-estimate 
# DALYs attributable to mental disorders.

# Draw in YLDs from Vigo et al.
data_rev$composite_ylds <-ifelse(data_rev$measure_id == 3, data_rev$rev_2016_value_agg, 0)

# Draw in YLLs from Walker et al.
data_rev$composite_ylls <-ifelse(data_rev$measure_id == 4, data_rev$rev_2015_value_agg, 0)

# Combine to calculate DALYs
data_rev_composite <- data_rev %>% select(location_id, numeric_name, paf_formula, measure_id, composite_ylds, composite_ylls) %>%
  group_by(location_id, numeric_name, paf_formula) %>%
  mutate(composite_dalys = sum(composite_ylds, composite_ylls))
data_rev_composite <- data_rev_composite %>% filter(measure_id == 2) %>% ungroup() %>% unique() %>% select("location_id", "numeric_name", "paf_formula", "composite_dalys") 
data_rev <- left_join(data_rev, data_rev_composite, by = c("location_id", "numeric_name", "paf_formula"))
data_rev$composite_dalys[data_rev$measure_id != 2] <- 0
data_rev$composite_ylds[data_rev$measure_id != 3] <- 0
data_rev$composite_ylls[data_rev$measure_id != 4] <- 0

# Draw in deaths from Walker et al.
data_rev$composite_deaths <-ifelse(data_rev$measure_id == 1, data_rev$rev_2015_value_agg, 0)
data_rev$composite_deaths[data_rev$measure_id != 1] <- 0

# Combined columns, clean up, re-organizing, tidying, and labeling
data_rev$composite <- data_rev$composite_ylds + data_rev$composite_ylls + data_rev$composite_dalys + data_rev$composite_deaths
data_rev <- data_rev %>% select(!c("composite_ylds":"composite_deaths")) %>%  relocate(original_value_agg, .after = paf_formula) %>%
  relocate(numeric_name, .after = prevalence) %>%  
  rename("Composite method" = composite,
         "PAF method" = rev_2015_value_agg,
         "2016 reallocation approach" = rev_2016_value_agg,
         "GBD 2019" = original_value_agg)

data_rev$cause_name <- "Mental disorders"
data_rev <- data_rev %>% gather("estimate", "number", "GBD 2019":"Composite method")

data_rev$estimate_id <- ifelse(data_rev$estimate == "Composite method", 4, 
                               ifelse(data_rev$estimate == "PAF method", 3,
                                      ifelse(data_rev$estimate == "2016 reallocation approach", 2, 1)))
data_rev <- data_rev %>% relocate(estimate_id, .after = estimate)

# Calculate rates and percent of burden
data_rev <- data_rev %>% relocate (population_100k, .after = population)
data_rev$rate_per_100k <- data_rev$number/data_rev$population_100k
data_rev$percent <- data_rev$number/data_rev$measure_total*100

############################
##        COSTING         ##
############################

# Import GDP
gdp_data <- read.csv(file = file.path(datapath,# "wdi_gdp_ppp.csv"  # Using PPP
                                      "wdi_gdp.csv"                  # Not using PPP
)) %>%
  select(-(c(3:62))) %>% 
  rename("gdp" = "X2019")
gdp <- gdp_data %>% select(c(iso_code, gdp))
data_rev <- left_join(data_rev, gdp, by = "iso_code")
data_rev <- data_rev %>% relocate(gdp, .after = iso_code)
data_rev$gdp[data_rev$location_id == 1] <- gdp_data$gdp[gdp_data$iso_code == "WLD"] # GDP for the world

# Calculate GDP per capita, with reference to IHME population values, for consistency
data_rev$gdp_per_capita <- data_rev$gdp/data_rev$population
data_rev <- data_rev %>% relocate (gdp_per_capita, .after = gdp)

# Clean up

# Import GDP PPP
gdp_data <- read.csv(file = file.path(datapath, "wdi_gdp_ppp.csv"  # Using PPP
                                      #"wdi_gdp.csv"                  # Not using PPP
)) %>%
  select(-(c(3:62))) %>% 
  rename("gdp_ppp" = "X2019")
gdp <- gdp_data %>% select(c(iso_code, gdp_ppp))
data_rev <- left_join(data_rev, gdp, by = "iso_code")
data_rev <- data_rev %>% relocate(gdp_ppp, .after = iso_code)
data_rev$gdp_ppp[data_rev$location_id == 1] <- gdp_data$gdp_ppp[gdp_data$iso_code == "WLD"] # GDP for the world

# Calculate GDP per capita, with reference to IHME population values, for consistency
data_rev$gdp_ppp_per_capita <- data_rev$gdp_ppp/data_rev$population
data_rev <- data_rev %>% relocate (gdp_ppp_per_capita, .after = gdp_ppp)

# Clean up

############################
# Cost per DALY

# Method 1: Copenhagen Consensus 1 - $1,000
data_rev$cost_cc1 <- ifelse(data_rev$measure_id == 2,
                            data_rev$number * 1000, 0)

# Method 2: Copenhagen Consensus 2 - $5,000
data_rev$cost_cc2 <- data_rev$cost_cc1 * 5

# Method 3: WHO - GDP/capita
data_rev$cost_who1 <-ifelse(data_rev$measure_id == 2,
                            data_rev$number * data_rev$gdp_per_capita, 0)

# Method 4: WHO - GDP/capita * 3
data_rev$cost_who2 <- data_rev$cost_who1 * 3

# Method 5: WHO - GDP PPP/capita
data_rev$cost_who3 <-ifelse(data_rev$measure_id == 2,
                            data_rev$number * data_rev$gdp_ppp_per_capita, 0)

# Method 6: WHO - GDP PPP/capita * 3
data_rev$cost_who4 <- data_rev$cost_who3 * 3

##################################
##         PAF FORMULA          ##
##################################

# Select which of the PAF formulae to use for the remainder of the analysis

data_rev <- data_rev %>% filter(paf_formula == "paf2") # Conventional formula
# data_rev <- data_rev %>% filter(paf_formula == "paf4") # Case prevalence formula

################################
##         MAP PREP           ##
################################

data_rev_gbd <- data_rev %>% filter(estimate_id == 1) %>% 
  select (c(measure_id, location_id, numeric_name, paf_formula, number:cost_who4)) %>%
  rename(
    "number_gbd" = "number",
    "rate_per_100k_gbd" = "rate_per_100k",
    "percent_gbd" = "percent",
    "cost_cc1_gbd" = "cost_cc1",
    "cost_cc2_gbd" = "cost_cc2",
    "cost_who1_gbd" = "cost_who1",
    "cost_who2_gbd" = "cost_who2",
    "cost_who3_gbd" = "cost_who3",
    "cost_who4_gbd" = "cost_who4")

data_rev <- left_join(data_rev, data_rev_gbd, by = c("location_id", "measure_id", "numeric_name", "paf_formula"))

data_rev <- data_rev %>% mutate(
  number_diff = number - number_gbd,
  rate_per_100k_diff = rate_per_100k - rate_per_100k_gbd,
  percent_diff = percent - percent_gbd,
  cost_cc1_diff = cost_cc1 - cost_cc1_gbd,
  cost_cc2_diff = cost_cc2 - cost_cc2_gbd,
  cost_who1_diff = cost_who1 - cost_who1_gbd,
  cost_who2_diff = cost_who2 - cost_who2_gbd,
  cost_who3_diff = cost_who3 - cost_who3_gbd,
  cost_who4_diff = cost_who4 - cost_who4_gbd)

# Import regional data
region <- read_excel(path = file.path(datapath, "regions.xlsx"))

# Merge
region <- region %>% select (iso_code, who_region, ihme_region)
data_rev <- full_join(data_rev, region, by = "iso_code")

# WHO region map prep

data_rev <- data_rev %>% group_by(estimate_id, measure_id, numeric_name, paf_formula, who_region) %>% 
  mutate(who_region_pop100k = sum(population_100k),
         who_region_number = sum(number),
         who_region_measure_total = sum(measure_total),
         who_region_gdp = sum(gdp, na.rm=TRUE),
         who_region_gdp_ppp = sum(gdp_ppp, na.rm=TRUE))%>%
  mutate(who_region_rate_per_100k = who_region_number/who_region_pop100k,
         who_region_percent = who_region_number/who_region_measure_total*100,
         who_region_population = who_region_pop100k * 100000) %>%
  mutate(who_region_gdp_per_capita = who_region_gdp/who_region_population,
         who_region_gdp_ppp_per_capita = who_region_gdp_ppp/who_region_population,
         who_region_cost_cc1 = ifelse(measure_id == 2,
                                      who_region_number * 1000, 0),
         who_region_cost_cc2 = who_region_cost_cc1 * 5,
         who_region_cost_who1 = ifelse(measure_id == 2,
                                       who_region_number * who_region_gdp_per_capita, 0),
         who_region_cost_who2 = who_region_cost_who1 * 3,
         who_region_cost_who3 = ifelse(measure_id == 2,
                                       who_region_number * who_region_gdp_ppp_per_capita, 0),
         who_region_cost_who4 = who_region_cost_who3 * 3) %>%
  ungroup()

data_rev_gbd_who_region<- data_rev %>% filter(estimate_id == 1) %>% 
  select (c(measure_id, location_id, numeric_name, paf_formula, who_region_number:who_region_cost_who4)) %>%
  rename(
    "who_region_number_gbd" = "who_region_number",
    "who_region_rate_per_100k_gbd" = "who_region_rate_per_100k",
    "who_region_percent_gbd" = "who_region_percent",
    "who_region_cost_cc1_gbd" = "who_region_cost_cc1",
    "who_region_cost_cc2_gbd" = "who_region_cost_cc2",
    "who_region_cost_who1_gbd" = "who_region_cost_who1",
    "who_region_cost_who2_gbd" = "who_region_cost_who2",
    "who_region_cost_who3_gbd" = "who_region_cost_who3",
    "who_region_cost_who4_gbd" = "who_region_cost_who4"
  )  %>%
  select(measure_id, location_id, numeric_name, paf_formula, who_region_number_gbd, who_region_rate_per_100k_gbd, who_region_percent_gbd,
         who_region_cost_cc1_gbd, who_region_cost_cc2_gbd, who_region_cost_who1_gbd, 
         who_region_cost_who2_gbd, who_region_cost_who3_gbd, who_region_cost_who4_gbd)

data_rev <- left_join(data_rev, data_rev_gbd_who_region, by = c("location_id", "measure_id", "numeric_name", "paf_formula"))

data_rev <- data_rev %>% mutate(
  who_region_number_diff = who_region_number - who_region_number_gbd,
  who_region_rate_per_100k_diff = who_region_rate_per_100k - who_region_rate_per_100k_gbd,
  who_region_percent_diff = who_region_percent - who_region_percent_gbd,
  who_region_cost_cc1_diff = who_region_cost_cc1 - who_region_cost_cc1_gbd,
  who_region_cost_cc2_diff = who_region_cost_cc2 - who_region_cost_cc2_gbd,
  who_region_cost_who1_diff = who_region_cost_who1 - who_region_cost_who1_gbd,
  who_region_cost_who2_diff = who_region_cost_who2 - who_region_cost_who2_gbd,
  who_region_cost_who3_diff = who_region_cost_who3 - who_region_cost_who3_gbd,
  who_region_cost_who4_diff = who_region_cost_who4 - who_region_cost_who4_gbd)

# IHME region  map prep

data_rev <- data_rev %>% group_by(estimate_id, measure_id, numeric_name, paf_formula, ihme_region) %>% 
  mutate(ihme_region_pop100k = sum(population_100k),
         ihme_region_number = sum(number),
         ihme_region_measure_total = sum(measure_total),
         ihme_region_gdp = sum(gdp, na.rm=TRUE),
         ihme_region_gdp_ppp = sum(gdp_ppp, na.rm=TRUE))%>%
  mutate(ihme_region_rate_per_100k = ihme_region_number/ihme_region_pop100k,
         ihme_region_percent = ihme_region_number/ihme_region_measure_total*100,
         ihme_region_population = ihme_region_pop100k * 100000) %>%
  mutate(ihme_region_gdp_per_capita = ihme_region_gdp/ihme_region_population,
         ihme_region_gdp_ppp_per_capita = ihme_region_gdp_ppp/ihme_region_population,
         ihme_region_cost_cc1 = ifelse(measure_id == 2,
                                       ihme_region_number * 1000, 0),
         ihme_region_cost_cc2 = ihme_region_cost_cc1 * 5,
         ihme_region_cost_who1 = ifelse(measure_id == 2,
                                        ihme_region_number * ihme_region_gdp_per_capita, 0),
         ihme_region_cost_who2 = ihme_region_cost_who1 * 3,
         ihme_region_cost_who3 = ifelse(measure_id == 2,
                                        ihme_region_number * ihme_region_gdp_ppp_per_capita, 0),
         ihme_region_cost_who4 = ihme_region_cost_who3 * 3) %>%
  ungroup()

data_rev_gbd_ihme_region<- data_rev %>% filter(estimate_id == 1) %>% 
  select (c(measure_id, location_id, numeric_name, paf_formula, ihme_region_number:ihme_region_cost_who4)) %>%
  rename(
    "ihme_region_number_gbd" = "ihme_region_number",
    "ihme_region_rate_per_100k_gbd" = "ihme_region_rate_per_100k",
    "ihme_region_percent_gbd" = "ihme_region_percent",
    "ihme_region_cost_cc1_gbd" = "ihme_region_cost_cc1",
    "ihme_region_cost_cc2_gbd" = "ihme_region_cost_cc2",
    "ihme_region_cost_who1_gbd" = "ihme_region_cost_who1",
    "ihme_region_cost_who2_gbd" = "ihme_region_cost_who2",
    "ihme_region_cost_who3_gbd" = "ihme_region_cost_who3",
    "ihme_region_cost_who4_gbd" = "ihme_region_cost_who4"
  )  %>%
  select(measure_id, location_id, numeric_name, paf_formula, ihme_region_number_gbd, ihme_region_rate_per_100k_gbd, ihme_region_percent_gbd,
         ihme_region_cost_cc1_gbd, ihme_region_cost_cc2_gbd, ihme_region_cost_who1_gbd, 
         ihme_region_cost_who2_gbd, ihme_region_cost_who3_gbd, ihme_region_cost_who4_gbd)

data_rev <- left_join(data_rev, data_rev_gbd_ihme_region, by = c("location_id", "measure_id", "numeric_name", "paf_formula"))

data_rev <- data_rev %>% mutate(
  ihme_region_number_diff = ihme_region_number - ihme_region_number_gbd,
  ihme_region_rate_per_100k_diff = ihme_region_rate_per_100k - ihme_region_rate_per_100k_gbd,
  ihme_region_percent_diff = ihme_region_percent - ihme_region_percent_gbd,
  ihme_region_cost_cc1_diff = ihme_region_cost_cc1 - ihme_region_cost_cc1_gbd,
  ihme_region_cost_cc2_diff = ihme_region_cost_cc2 - ihme_region_cost_cc2_gbd,
  ihme_region_cost_who1_diff = ihme_region_cost_who1 - ihme_region_cost_who1_gbd,
  ihme_region_cost_who2_diff = ihme_region_cost_who2 - ihme_region_cost_who2_gbd,
  ihme_region_cost_who3_diff = ihme_region_cost_who3 - ihme_region_cost_who3_gbd,
  ihme_region_cost_who4_diff = ihme_region_cost_who4 - ihme_region_cost_who4_gbd)

# Import map data
world <- ne_countries(scale = "medium", returnclass = "sf")
world$iso_code <- world$gu_a3
world$iso_code[world$adm0_a3 == "SDS"] <- "SSD"

# Re-order estimates: original, Vigo, Walker, composite

data_rev <- data_rev %>% 
  mutate(estimate_newlab = recode(estimate, 
                                  'GBD 2019' = "Original approach",                
                                  '2016 reallocation approach' = "Reallocation approach",
                                  'PAF method' = "PAF method",              
                                  'Composite method'= "Composite approach"  
  ))

data_rev$estimate_newlab <- factor(data_rev$estimate_newlab ,      # Reordering group factor levels
                                   levels = c("Original approach",
                                              "Reallocation approach",
                                              "PAF method",
                                              "Composite approach"))
# Merging map and dataframe
data_rev_map <- full_join(world, data_rev, by = "iso_code") %>% filter(numeric_name == "val")

# Clean up

# Standard subtitles, and captions
subtitle_1 <- "Value per DALY: $1,000"
subtitle_2 <- "Value per DALY: $5,000"
subtitle_3 <- "Value per DALY: GDP/capita"
subtitle_4 <- "Value per DALY: 3 X GDP/capita"
#caption <- "Source: Global Burden of Disease Study. Reallocation approach from Vigo et al. 2016."

```

```{r paper1code_results}

####################################
##       REGIONAL CHARTS          ##
####################################

who_region_positions <- c("African Region", "Eastern Mediterranean Region", "European Region",
                          "Region of the Americas", "South-East Asia Region", "Western Pacific Region",  "#NA")

who_region_pal <-  c("#E74C3C", "#E67E22" , "#F4D03F", "#45B39D", "#5499C7", "#AF7AC5")

ihme_region_positions <- c("East Asia", "Southeast Asia", "Oceania", "Central Asia", "Eastern Europe", "Central Europe",
                           "Caribbean", "Central Latin America", "Tropical Latin America", "Andean Latin America", "North Africa and Middle East",
                           "Southern Sub-Saharan Africa", "Western Sub-Saharan Africa", "Central Sub-Saharan Africa", "Eastern Sub-Saharan Africa",
                           "South Asia", "Southern Latin America", "Western Europe", "High-income North America",
                           "Australasia", "High-income Asia Pacific", "#NA")

ihme_pal_new <- c("#002673", "#005CE6", "#73DFFF", "#E54800", "#E69800", "#FFD37F",
                  "#216F00", "#33A600", "#4CE600", "#A3FF73", "#4E4E4E", 
                  "#730000", "#FF0000", "#FF7F7F", "#FFBEBE", 
                  "#E6E600", "#73004C", "#A80084", "#C500FF", "#E600A9", "#FFBEE8", "#7F7F7F")

ihme_region_pal <- c("#8C95F1", "#D7ECFC" , "#B6CDD4", "#EAA144", "#F3DA89", "#F9EFC2", "#B1F585", "#B6F2D5",
                     "#E3FD8C", "#E2FDD7", "#99ACA4", "#9D5F56", "#E989A9", "#F5DBD8", "#F8EEEF",
                     "#F4FF7C", "#C42AF1", "#E986F8", "#D7C2FC", "#DD9BE9", "#FAE5FE", "#7F7F7F")

############################
##        TABLES          ##
############################

# Table 1: Burden of disease due to mental disorders

data_global <- data_rev %>% filter(location_id == 1)  %>% filter(estimate_id != 3)  %>% select(location_name, measure_name, numeric_name, estimate, number, percent) %>%
  mutate(number = round(number/1000000, 1),
         percent = round(percent, 1))

data_regional <- data_rev  %>% filter(estimate_id != 3)  %>% 
  group_by(ihme_region,measure_name, numeric_name, estimate) %>% mutate(region_total = sum(measure_total),
                                                                        region_number = sum(number)) %>%
  mutate(region_percent = round(region_number / region_total * 100, 1)) %>% select(ihme_region, measure_name, numeric_name, estimate, region_number, region_percent) %>%
  rename("number" = "region_number",
         "percent" = "region_percent",
         "location_name" = "ihme_region") %>%
  ungroup() %>%
  unique() %>% filter(!is.na(location_name)) %>%
  mutate(number = round(number/1000000, 1)) %>% arrange((location_name))

data_income <- data_rev  %>% filter(estimate_id != 3)  %>% 
  group_by(income_level,measure_name, numeric_name, estimate) %>% mutate(region_total = sum(measure_total),
                                                                         region_number = sum(number)) %>%
  mutate(region_percent = round(region_number / region_total * 100, 1)) %>% select(income_level, measure_name, numeric_name, estimate, region_number, region_percent) %>%
  rename("number" = "region_number",
         "percent" = "region_percent",
         "location_name" = "income_level") %>%
  ungroup() %>%
  unique() %>% filter(!is.na(location_name)) %>%
  mutate(number = round(number/1000000, 1))

location_name <- c("H","UM","LM","L")
rank <- c(1,2,3,4)
fullname <- c("High income",
              "Upper-middle income",
              "Lower-middle income",
              "Low income")

incomelevels <- data.frame(location_name,rank, fullname)

data_income <- left_join(data_income, incomelevels, by = "location_name") %>% arrange(rank) %>% select(!c(location_name, rank)) %>% rename("location_name" = "fullname")

data_table1 <- rbind(data_global, data_income, data_regional) %>% 
  gather("metric","point","number":"percent" ) %>%  
  pivot_wider(names_from = c(numeric_name, metric, estimate), values_from = point)

col_order <- c(
  "measure_name",
  "location_name",                                    
  "val_number_GBD 2019",                             
  "lower_number_GBD 2019",                            
  "upper_number_GBD 2019",   
  "val_percent_GBD 2019",                            
  "lower_percent_GBD 2019",  
  "upper_percent_GBD 2019",                           
  "val_number_2016 reallocation approach",   
  "lower_number_2016 reallocation approach",  
  "upper_number_2016 reallocation approach",  
  "val_percent_2016 reallocation approach",  
  "lower_percent_2016 reallocation approach", 
  "upper_percent_2016 reallocation approach", 
  "val_number_Composite method",           
  "lower_number_Composite method",          
  "upper_number_Composite method",          
  "val_percent_Composite method",          
  "lower_percent_Composite method",  
  "upper_percent_Composite method")

data_table1 <- data_table1[, col_order]

data_table1 <- data_table1 %>% arrange(measure_name)

write.csv(data_table1, file = paste0(resultspath, "/table1.csv"))

# Table 2: Value of economic welfare estimates

data_global <- data_rev %>% 
  filter(location_id == 1)  %>% 
  filter(measure_id == 2) %>%
  filter(estimate_id != 3)  %>% 
  select(location_name, numeric_name, 
         estimate, cost_cc1, cost_cc2, cost_who1, cost_who2, cost_who3, cost_who4)

data_global$location_name <- "Global"

data_regional <- data_rev  %>%   
  filter(measure_id == 2) %>%
  filter(estimate_id != 3)  %>% 
  select(ihme_region,numeric_name, estimate, ihme_region_cost_cc1,
         ihme_region_cost_cc2, ihme_region_cost_who1, ihme_region_cost_who2, ihme_region_cost_who3, ihme_region_cost_who4) %>%
  unique() %>%
  rename("location_name" =ihme_region,
         "cost_cc1" = ihme_region_cost_cc1,
         "cost_cc2" = ihme_region_cost_cc2,
         "cost_who1" = ihme_region_cost_who1,
         "cost_who2" = ihme_region_cost_who2,
         "cost_who3" = ihme_region_cost_who3,
         "cost_who4" = ihme_region_cost_who4) %>%
  filter(!is.na(location_name)) %>% 
  arrange((location_name))

data_income <- data_rev  %>% 
  filter(measure_id == 2) %>%
  filter(estimate_id != 3)  %>%  
  select(income_level, estimate, numeric_name, cost_cc1, cost_cc2, cost_who1, cost_who2,  cost_who3, cost_who4) %>%
  pivot_longer(cols = cost_cc1:cost_who4) %>%
  group_by(income_level,numeric_name, name, estimate) %>% 
  mutate(cost_total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  select(income_level, numeric_name, name, cost_total, estimate) %>% unique() %>%
  rename("location_name" = income_level)

location_name <- c("H","UM","LM","L")
rank <- c(1,2,3,4)
fullname <- c("High income",
              "Upper-middle income",
              "Lower-middle income",
              "Low income")

incomelevels <- data.frame(location_name,rank, fullname)

data_income <- left_join(data_income, incomelevels, by = "location_name") %>% 
  arrange(rank) %>% select(!c(location_name, rank)) %>% rename("location_name" = "fullname")

data_income <- data_income %>% 
  pivot_wider(values_from = cost_total, names_from = name) %>%
  filter(!is.na(location_name))

#### Pivot wider, change column names, drop NA for income levels

data_table2 <- rbind(data_global, data_income, data_regional) %>% 
  arrange(factor(numeric_name, levels = c("val", "lower", "upper"))) %>%
  arrange(factor(estimate, levels = c("GBD 2019", "2016 reallocation approach", "Composite method"))) %>%
  pivot_longer(cols = cost_cc1:cost_who4) %>%
  mutate(value = round(value/1000000000000,2)) %>%
  pivot_wider(names_from = c(numeric_name, estimate), values_from = value) %>%
  arrange(name) %>%
  relocate(name, .before = "location_name")

# write.csv(data_table2, file = paste0(resultspath, "/table2.csv"))

# Appendix 2: Value of economic welfare estimates (CC)

appendix_table <- data_rev %>% filter(location_id == 1) %>%  
  filter(measure_id == 2) %>% filter(estimate_id != 3) %>% 
  select(estimate, numeric_name, cost_cc1, cost_cc2, cost_who1, cost_who2, cost_who3, cost_who4)
appendix_table$cost_cc1 <- round(appendix_table$cost_cc1/1000000000000,2)
appendix_table$cost_cc2 <- round(appendix_table$cost_cc2/1000000000000,2)
appendix_table$cost_who1 <- round(appendix_table$cost_who1/1000000000000,2)
appendix_table$cost_who2 <- round(appendix_table$cost_who2/1000000000000,2)
appendix_table$cost_who3 <- round(appendix_table$cost_who3/1000000000000,2)
appendix_table$cost_who4 <- round(appendix_table$cost_who4/1000000000000,2)
appendix_table <- appendix_table %>% select(!c(cost_who1, cost_who2)) %>% arrange(estimate, factor(numeric_name, levels = c("val", "lower", "upper")))
appendix_table_t <- transpose(appendix_table)
